---
title: "simpleRegress"
author: "Wilhelmus Elersich"
date: "Fall 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```
```{r}
plot(WEIGHT~SEX,data=bears)
plot(LENGTH~SEX, data=bears)
```

We can see that, from these plots, that the sex value of 1 corresponds to male bears, and the sex value of 2 corresponds to female bears.

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
It would appear that neck circumference is positively correlated with body length in bears.
## now predict the length of a bear who has a neck of 17, 22 and 26
By looking at plot, we can predict a bear with a neck of 17 having a length of 50, a bear with a neck of 22 having a length of 60, and a bear with a neck of 26 having a length of 70. But a computer can predict better than a human can.

```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```


## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).
```{r}
LMHEADLEN <- lm(LENGTH ~ HEADLEN, data=bears)
summary(LMHEADLEN)
```
```{r}
LMCHEST <- lm(LENGTH ~ CHEST, data=bears)
summary(LMCHEST)
```
```{R}
LMNECK <- lm(LENGTH ~ NECK, data=bears)
summary(LMNECK)
```
```{r}
LMWEIGHT <- lm(LENGTH ~ WEIGHT, data=bears)
summary(LMWEIGHT)
```
```{r}
LMHEADWTH <- lm(LENGTH ~ HEADWTH, data=bears)
summary(LMHEADWTH)
```
```{r}
LMAGE <- lm(LENGTH ~ AGE, data=bears)
summary(LMAGE)
```

(IN DESCENDING ORDER OF PERFORMANCE)  
1. HEADLEN  
2. CHEST  
3. NECK  
4. WEIGHT  
5. HEADWTH  
6. AGE  

### For male bears:
```{R}
mbears <-  filter(bears, SEX==1)
```
```{R}
LMNECK <- lm(LENGTH ~ NECK, data=mbears)
summary(LMNECK)
```
```{r}
LMHEADLEN <- lm(LENGTH ~ HEADLEN, data=mbears)
summary(LMHEADLEN)
```
```{r}
LMCHEST <- lm(LENGTH ~ CHEST, data=mbears)
summary(LMCHEST)
```
```{r}
LMWEIGHT <- lm(LENGTH ~ WEIGHT, data=mbears)
summary(LMWEIGHT)
```
```{r}
LMHEADWTH <- lm(LENGTH ~ HEADWTH, data=mbears)
summary(LMHEADWTH)
```
```{r}
LMAGE <- lm(LENGTH ~ AGE, data=mbears)
summary(LMAGE)
```

(IN DESCENDING ORDER OF PERFORMANCE)  
1. NECK  
2. HEADLEN  
3. CHEST  
4. WEIGHT  
5. HEADWTH  
6. AGE  

There is a difference between the results expected for male bears, and bears in general.
